MR-PFU: A suite of R packages for energy conversion chain analysis

### 2023-07-28 paper.Rmd
More edits.
Added acknowledgements.

Further development followed from selecting the PSUT framework.
First, the RUVY matrices carry the challenge that 
different countries and years 
have varying energy carriers (products) and
varying energy conversion machines (industries),
meaning that RUVY matrices 
have differing row names, differing column names, and differing sizes.
To get around this challenge, we created the
[matsbyname][link-matsbyname] package [@Heun-matsbyname:2023]
which enables matrix mathematics 
that respects row and column names, 
inserting **0** vectors when needed.
Second,
wanted to be able to perform _matrix_ mathematics 
as easily as _scalar_ mathematics
in `R` data frames
using the [tidyverse][link-tidyverse] syntax [@Wickham:2019]. 
We developed the [matsindf][link-matsindf] package [@Heun-matsindf:2023]
to enable this functionality.
Finally, manipulating row and column names proved to be a challenge, 
especially for RUVY matrices in [matsindf][link-matsindf] data frames, so 
we developed the [RCLabels][link-rclabels] package [@Heun-RCLabels:2023]
to assist.
The table below summarizes these packages, 
all of which are generally useful and available on 
[CRAN][link-cran].

@@

MKH acknowledges the support of Calvin University for 
a sabbatical during which he developed the PSUT framework
and course releases during which he wrote many of the `R` packages
described in this paper.
ZM acknowledges ******. 
EA acknowledges ******.
PEB acknowledges ******.

>>>>

Further development followed selection of the PSUT framework.
First, the RUVY matrices carry the challenge that 
different countries and years 
have varying energy carriers (products) and
varying energy conversion machines (industries),
meaning that RUVY matrices 
have differing row names, differing column names, and differing sizes.
To get around this challenge, we created the
[matsbyname][link-matsbyname] package [@Heun-matsbyname:2023]
which enables matrix mathematics 
that respects row and column names, 
inserting **0** vectors when needed.
Second,
wanted to be able to perform _matrix_ mathematics 
as easily as _scalar_ mathematics
in `R` data frames
using the [tidyverse][link-tidyverse] syntax [@Wickham:2019]. 
We developed the [matsindf][link-matsindf] package [@Heun-matsindf:2023]
to enable this functionality.
Finally, manipulating row and column names proved to be a challenge, 
especially for RUVY matrices in [matsindf][link-matsindf] data frames, so 
we developed the [RCLabels][link-rclabels] package [@Heun-RCLabels:2023]
to assist.
The table below summarizes these packages, 
all of which are generally useful and available on 
[CRAN][link-cran].

@@

MKH acknowledges the support of Calvin University for 
a sabbatical during which he developed the PSUT framework
and course releases during which he wrote many of the `R` packages
described in this paper.
ZM's and PEB's time was funded by 
the UK Research Council under EPSRC Fellowship award EP/R024254/1.
EA's time was funded by 
the University of Leeds (School of Earth and Environment).
